Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-02-11 ThesisJP/Background/background.tex
Updated cpu vs gpu figure.
Write on background.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{img/free_lunsh.png}
\caption{Illustration of transitions between different eras in the computer industry. Currently there are three ongoing transitions, multi-core, heterogeneous-core and cloud-core. Image courtesy of Herb Sutter (herbsutter.com).}
\label{fig:jungle}
\end{figure}

Around the same time as CPU manufactures hit the frequency wall, people started experimenting with programmable shaders which recently had been introduced in the field of graphics programming. The graphics processing pipeline had before this been a fixed function pipeline, where e.g. geometry and textures where fed to the graphics processing unit (GPU) in one end and a rasterized rendered image was outputted in the other. What programmable shaders  introduced was the ability to replace some steps in the fixed-function pipeline with custom computations. This made it possible to utilize the GPU for other tasks than just rendering graphics \cite{Seland2007}. The rationale behind this exploration was that hidden in the GPU was processing powers an order of magnitude larger than what the CPU could provide at the same time (See Fig.\,\ref{fig:cpu_vs_gpu}). Actually, at this time the GPU had already gone multi-core, driven by an ever-increasing demand for more realistic computer games. What also made this possible was the inherent parallel nature of the rendering problem. When geometry is projected and rasterized the same instruction is typically needed for a lot of data at the same time. This makes a special kind of instruction, Single Instruction Multiple Data (\nom{SIMD}{Single instruction multiple data (instruction)}) instruction, especially suited for this kind of problem. The rendering pipeline also exhibits minimal execution path branching, and data are typically used once. Designers of GPUs could therefore skip advance feature, found in most CPUs, like branch prediction and different levels of data caching. This reduced the silicon footprint of each core and made it possible to add multiple SIMD cores to the same chip before CPU designers where able to do the same. Today, even if more advanced caching has added to GPUs, this is still the main reasons why GPUs have higher peak performance than CPUs (as shown in Fig.\,\ref{fig:cpu_vs_gpu}).

The first adopters of GPU programming, or General-purpose programming on graphics processing units (GPGPU), wanted to use the GPU to accelerate computationally demanding algorithms.

@@

Three eras at the same time: Multi-core, Heterogeneous-cores, and Cloud-cores.

\subsection{Comparing CPU and GPU performance} 
It is common to say that a CPU has few but advance generic cores, and the GPU has many but simple cores.

>>>>

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{img/free_lunsh.png}
\caption{Illustration of transitions between different eras in the computer industry. Currently there are three ongoing transitions; multi-core, heterogeneous-core and cloud-core. Image courtesy of Herb Sutter (herbsutter.com).}
\label{fig:jungle}
\end{figure}

Around the same time as CPU manufactures hit the frequency wall, people started experimenting with programmable shaders which recently had been introduced in the field of graphics programming. The graphics processing pipeline had before this been a fixed function pipeline, where e.g. geometry and textures where fed to the graphics processing unit (GPU) in one end and a rasterized rendered image was outputted in the other. What programmable shaders  introduced was the ability to replace some steps in the fixed-function pipeline with custom computations. This made it possible to utilize the GPU for other tasks than just rendering graphics \cite{Seland2007}. The rationale behind this exploration was that GPUs were already a highly available processor, and hidden behind its graphics interface was processing powers an order of magnitude larger than what the CPU could provide (Fig.\,\ref{fig:cpu_vs_gpu}). Actually, at the time when CPUs hit the frequency wall the GPU had already gone multi-core, driven by an ever-increasing demand for more realistic computer games. The rendering pipeline exhibits minimal execution path branching, and data are typically used once. Designers of GPUs could therefore skip advance feature, found in most CPUs, like branch prediction and different levels of data caching. This reduced the silicon footprint of each core and made it possible to add multiple cores to the same chip before CPU designers where able to do the same. Today, even if more advanced caching has been added to GPUs, this is still the main reason why GPUs have higher peak performance than CPUs (Fig.\,\ref{fig:cpu_vs_gpu}). Another reason is the inherent parallel nature of the rendering problem. When geometry is projected and rasterized the same instruction is typically needed for a lot of data at the same time. This makes a special kind of instruction, Single Instruction Multiple Data (\nom{SIMD}{Single instruction multiple data (instruction)}) instruction, especially suited for this kind of problem. The differences between how SIMD instructions are utilized in modern CPUs and GPUs is discussed in Section \ref{sec:cpu_vs_gpu}.

For the first adopters of GPU programming, or General-purpose programming on graphics processing units (GPGPU) as it was named, it soon became evident that offloading all types of computations to the GPU was not always a good idea. In most cases it is important to balance the load equally between the CPU and GPU to obtain their combined computationally power. Balancing typically means that highly parallel tasks should be offloaded to the GPU and serialized work should stay on the CPU. This paradigm of utilizing specialized cores for solving specific problems is today known as heterogeneous computing. The specialized core can

@@

Three currently ongoing transitions: Multi-core, Heterogeneous-cores, and Cloud-cores.

This thesis is focusing on heterogeneous computing with CPUs and GPUs.

\subsection{Comparing CPU and GPU performance}\label{sec:cpu_vs_gpu}
It is common to say that a CPU has few but advance generic cores, and the GPU has many but simple cores.
